---
title: "Global Renewable Energy Leaders in 2023"
author: Master
format:
  html:
    toc: true
    toc-depth: 2
editor: visual
---

## Top 10 Countries by Renewable Energy Share in 2023

The dataset includes annual values for each country from 1965 to 2023, showing the percentage of total primary energy derived from renewable sources.

The output is shown in **Table 1**, which lists the countries with the highest renewable energy share in 2023. In addition, **Figure 1** provides a horizontal bar chart to visually compare the leading countries.

```{r filter-top10, message=FALSE, warning=FALSE, echo=FALSE}
library(readr)
library(tidyverse)
data <- read_csv("renewable-share-energy.csv/renewable-share-energy.csv")

top10_clean <- data %>%
  filter(Year == 2023, 
         !Entity %in% c("World", "Africa", "Asia", "Europe", "Oceania", 
                        "North America", "South America", "Eastern Africa (EI)")) %>%
  arrange(desc(`Renewables (% equivalent primary energy)`)) %>%
  slice(1:10)

```

```{r table-top10, echo=FALSE, message=FALSE, warning=FALSE}
library(knitr)

top10_clean %>%
  select(Entity, Code, Year, `Renewables (% equivalent primary energy)`) %>%
  rename(
    Country = Entity,
    `Renewables (%)` = `Renewables (% equivalent primary energy)`
  ) %>%
  kable(caption = "Table 1. Top 10 Countries by Share of Renewable Energy in 2023")
```

```{r fig-top10, fig.cap="Share of Renewable Energy in Top 10 Countries (2023)", echo=FALSE, message=FALSE, warning=FALSE}

ggplot(top10_clean, aes(x = reorder(Entity, `Renewables (% equivalent primary energy)`), 
                  y = `Renewables (% equivalent primary energy)`)) +
  geom_col(fill = "steelblue") +
  coord_flip() +
  labs(
    title = "Top 10 Countries by Renewable Energy Share (2023)",
    x = "Country",
    y = "Renewables (% of Primary Energy)"
  ) +
  theme_minimal()

```
